Constraint Grammar as a SAT problem

### 2015-03-30 paper/nodalida2015.tex
more stuff

\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\title{Constraint Grammar as a SAT problem}

@@

\noindent These three features might change the way that rules are written. 
We predict that less rules are needed; whether this holds in
the order of thousands of rules remains to be tested.
Getting rid of ordering could ease the task of the grammar writer,
removing the responsibility to estimate the best sequence of rule applications.
However, the implementation should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.

@@

We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
Treating the chosen tagset as the gold standard, the disambiguation by
SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
Similar patterns were observed with small (\textless{}20) rule sets written by
the authors; depending on the subset, SATCG and VISL CG-3 had
a difference of at most 1.5 \%. Introducing rules one by one up to 19, the
performance improved in a very similar rate, with less than 0.5 \%
difference between implementations at each new rule.

Additional tests could include plugging SATCG into Apertium
translation pipeline, and comparing the translation quality.

>>>>

\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) --
  (.25,.15) -- cycle;}
\newcommand{\Mypm}{\mathbin{\tikz [x=1.4ex,y=1.4ex,line width=.1ex] \draw (0.0,0) -- (1.0,0) (0.5,0.08) -- (0.5,0.92) (0.0,0.5) -- (1.0,0.5);}}%

\title{Constraint Grammar as a SAT problem}

@@

\noindent These three features might change the way that rules are written. 
We predict that less rules are needed; whether this holds in
the order of thousands of rules remains to be tested.
Getting rid of ordering could ease the task of the grammar writer,
removing the responsibility to estimate the best sequence of rule applications.
However, the implementation should not be too different from the
existing ones---our preliminary evaluation against VISL CG-3 shows
promising results in this regard.
\pagebreak

@@

We took a tagged corpus\footnote{https://svn.code.sf.net/p/apertium/svn/branches/apertium-swpost/apertium-en-es/es-tagger-data/es.tagged} of 21865 words for Spanish, 
and a small constraint grammar\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}, produced independently of the authors.
% We ignored substitute rules, but kept subreadings and unification, even though our implementation doesn't handle them.
We only took select and remove rules (no substitute/iff), after which the constraint grammar had 261 rules.
With this setup, we took the text of the tagged corpus with all ambiguities, and ran both VISL CG-3 and our implementation with the same grammar. 
Treating the chosen tagset as the gold standard, the disambiguation by
SATCG achieves 75.09 \% correct result and VISL CG-3 75.80 \%.
Similar patterns were observed with small (\textless{}20) rule sets written by
the authors; depending on the subset, SATCG and VISL CG-3 had
a difference of at most $\Mypm$ 1.5 \%. 
Introducing rules one by one up to 19, the
performance improved in a very similar rate, with less than 0.5 \%
difference between the systems at each new rule.

% Additional tests could include plugging SATCG into Apertium
% translation pipeline, and comparing the translation quality.
